% !TEX options=--shell-escape
\title{Zusammenfasung - Physik 1}
\author{
        Pascal Müller [pamuelle@student.ethz.ch]
}
\date{\today}

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[parfill]{parskip}
\usepackage{amsmath}
\usepackage{hyperref}

\title{Classification and Parameter Estimation of Gravitational Waves from
Binary-Black-Hole, Binary-Neutron-Star and Black-Hole-Neutron-Star Mergers in
Real LIGO Data using Deep Learning}

\begin{document}
\maketitle

\paragraph{Problem:}
Gravitational wave (GW) detection currently employs computationally expensive
methods for GW signal classification and parameter estimation. This leads to
delayed public alerts for new detections, making prompt subsequent
multimessenger follow-ups unfeasible. \\

The need for real-time detection of GW signals led to an increase in the
application of deep learning (DL) methods to classify GW signals and estimate
its parameters. Currently all detected GW signals are sourced by compact binary
coalescences (CBC). In particular: Binary neutron stars (BNS), binary black
holes (BBH) and neutron star - black hole (BHNS) merger. \\

George and Huerta showed \cite{BBHmerger} that not only can Convolutional Neural
Networks (CNN) be used to detect and estimate the parameters of GW signals from
BBH mergers in real LIGO noise (including real glitches) but also that CNNs
achieve performance comparable to matched-filtering methods. Palmen G. Krastev
showed \cite{NSmerger} that the same can be done for GW signals from BNS merger.
\\

Similar work for BHNS merger as well as research using a unified DL framework
for analyzing BNS and BHNS merger simultaneously is said to be conducted by
Palmen G. Krastev \cite{NSmerger} in future works.

\paragraph{Task:}
The task of this thesis is to create a synergy between GW physics and deep
learning. In particular the task will be to classify and estimate the parameters
\footnote{component masses ($m_1, m_2$), spin $(\mathbf{\hat{s}_1},
\mathbf{\hat{s}_2})$ and orbital eccentricity $e$} of GW signals from BBH, BNS
and  and BHNS mergers in real LIGO data using a (deep) neural network approach.
The focus will be on real-world application i.e. the neural network should also
correctly classify glitches as noise. Furthermore the possibility and usefulness
of a unified deep learning framework should also be researched. \\

The estimation of additional parameters such as for example the sky location,
inclination or tidal deformabilities might also pose as an additional research
goal.

\begin{thebibliography}{9}
\bibitem{NSmerger}
Plamen G. Krastev, Kiranjyot Gill, V. Ashley Villar, Edo Berger.
\href{https://arxiv.org/abs/2012.13101}
{\textit{Detection and Parameter Estimation of Gravitational Waves from Binary
Neutron-Star Mergers in Real LIGO Data using Deep Learning}}

\bibitem{BBHmerger}
Daniel George, E.A. Huerta.
\href{https://www.sciencedirect.com/science/article/pii/S0370269317310390}
{\textit{Deep Learning for real-time gravitational wave detection and parameter
estimation: Results with Advanced LIGO data}}
\end{thebibliography}

\end{document}
